---
title: debugging \LaTeX{} is very very frustrating
mainfont: Arial
mathfont: "Fira Math Regular"
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{makecell}
  - \usepackage{xcolor}
---

#####  TFC defined by increasing intergranular void ratio

A second mathematical means to define the TFC is the intergranular void ratio $e_{sand}$ [@Thevanayagam1998].
This parameter reflects the relative density of the larger grains.
Equation \@ref(eq:intergranular-void-ratio-equation) demonstrates that $e_{sand}$ is derived from phase relations by considering the solid fines as voids:

\begin{align}
e_{sand} &= \frac{V_{voids}}{V_{sand}} \nonumber \\
e_{sand} &= \frac{V_{clay} + V_{water}}{V_{sand}} \nonumber \\
e_{sand} &= \frac{\frac{m_{clay}}{G_{clay}} + \frac{m_{water}}{Gwater}}{\frac{m_{sand}}{G_{sand}}}; \quad m_{water} = w \cdot m_{solids} = w \cdot 1 = w \nonumber \\
e_{sand} &= \frac{\frac{m_{clay}}{G_{clay}} + \frac{{w}}{Gwater}}{\frac{m_{sand}}{G_{sand}}}
(\#eq:intergranular-void-ratio-equation)
\end{align}

Where $G$ represents the specific gravity of a phase and $m$ represents its mass fraction, with $m_{solids} = 1$.

The state of the sand grains can be considered in terms of their porosity rather than their void ratio:

\begin{align}
n_{sa}&=\frac{\frac{\left( m_{c_{mass-based}} \times  \rho_{bulk} \right)}{G_c}+\left( 1 - \frac{\rho_{bulk}}{G_s} \right)}{1}
(\#eq:lit-review-intergranular-porosity-equation)
\end{align}

A more complete derivation of Equation \@ref(eq:lit-review-intergranular-porosity-equation) is given in the Appendix (Section \@ref(generalized-intergranular-porosity-equation)).


```{r compute-phase-volumes}
source(here::here('supplemental_R_scripts/proposal-fig-generation/tfc-drawings/drawing-intergranular-void-ratio-75-pct-sand.R'))

# the script also calculates some values to use dynamically below.

```


```{r}

rounded_phase_volumes <- phase_volumes %>%
  dplyr::mutate(dplyr::across(.fns = round, digits =2))


rounded_phase_masses <- phase_masses %>%
  dplyr::mutate(dplyr::across(.fns = round, digits =2))


```

Figure \@ref(fig:intergranular-void-ratio-diagram) illustrates phase volumes for a hypthetical sand-clay mixture having transitional fines content $TFC=0.25$, i.e. `r 100*rounded_phase_masses$m_sa`% sand by mass. <!-- and $w=$ `r round(phase_diagram_w, digits=3)`.  -->
Assuming full saturation, $V_{sand}=$ `r rounded_phase_volumes$v_sa`, $V_{clay}=$ `r rounded_phase_volumes$v_c`, and $V_{water}=\theta=$ `r rounded_phase_volumes$v_w`. Total volume $V_{total}=1$. $e_{sand}$ is `r round((phase_volumes$v_w + phase_volumes$v_c)/(phase_volumes$v_sa), digits =2)` and $n_{sand}=$ `r round((phase_volumes$v_w + phase_volumes$v_c)/(phase_volumes$v_w + phase_volumes$v_c + phase_volumes$v_sa), digits = 2)`.
These are similar to typical $e_{min}$ or $n_{max}$ values for a pure, uniform sand.

```{r intergranular-void-ratio-diagram, fig.cap='The solid fines are considered as part of the voids phase when computing $e_{SA}$.', fig.scap='Intergranular void ratio of a sand-clay mixture.', out.width='60%', fig.pos='hptb'}
knitr::include_graphics(here::here('figs/pdf/intergranular-void-ratio-75-percent-sand.pdf'))

```

@Thevanayagam1998 proposed that $e_{sand}$ at a given $m_{clay}$ and $w$ be compared with the minimum void ratio of the pure "host" sand $e_{max_{pure~sand}}$.
When $e_{sand}$ exceeds $e_{max_{pure~sand}}$, the soil's behavior will no longer be governed by contacts between coarse grains because of their loose configuration.
Therefore @Thevanayagam1998 defined the TFC as the fines content where $e_{sand}=e_{max_{pure~sand}}$


##### Effect of coarse/fine particle size ratio {#tfc-size-ratio}

The solutions described above assume that the fines are infinitely small, fitting completely within the larger grains.
In reality, lodging of fines between the coarse grains often creates additional voids.
@Lade1998 defined a size ratio $R$ as the quotient of each component's mean grain size:

\begin{equation}
R=\frac{D_{50~sand}}{D_{50~fines}} \\
(\#eq:tfc-size-ratio)
\end{equation}


The disparity between calculations and experiments increases with decreasing size ratio [@Lade1998; @Zuo2015].
@Lade1998 concluded that when $R$ < 7, any increase in the size ratio imparts a relatively steep increase in packing density.
For further increases in $R$, the observed void ratio slowly decreases toward its theoretical minimum (Figure \@ref(fig:lade-1998-figure-7)).

```{r lade-1998-figure-7, fig.cap='When the ratio of particle diameters is < 7, increases in the size ratio improve packing density. Above R of 7-10, little density increase is observed. Reproduced from Lade et al. (1998).', fig.scap='Importance of particle diameter ratio for intergranular packing'}
knitr::include_graphics(here::here('images/illustrations/Lade_1998_Figure_7/Lade_1998_Figure_7.pdf'))
```

@Lade1998 also showed that a void between closely packed spheres of equal sizes has a diameter of ~ $\frac{1}{6.4}$ that of the large particles.
Therefore, there is a fundamental reason for the abrupt change in slope at $R$~7.

Numerical simulations indicate the intergranular void ratio is practically independent of $D_{50~fines}$ when $R$ > 100 [@Ueda2011].

The geometrical calculations by @Lade1998 assume a single particle size for each mix component.
The solid volume $V_s$ of a two-component mixture is further increased when each individual component is well-graded.
@Reed1995 provided solutions for the optimum ratios of multi-phase mixtures when each has a log-normal particle size curve.